
# 1 Libraries
```{r}
library(biomod2)
library(ggplot2)
library(gridExtra)
library(raster)
library(rasterVis)
library(sf)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidygeocoder)
library(tidyr)
library(tidyterra)
library(ggtext)
```

# 2 Environmental variables 
```{r}
variables <- raster :: stack (c(
  other_agri_dist =
    "~/model/ascii/agr_d_train.asc",
  con_dist = 
    "~/model/ascii/con_d_train.asc",
  decid_dist = 
    "~/model/ascii/dec_d_train.asc",
  lowcover_dist = 
    "~/model/ascii/low_d_train.asc", 
  mais_dist = 
    "~/model/ascii/maize_d_train.asc",
  water_dist = 
    "~/model/ascii/water_d_train.asc", 
  conif_frac = 
    "~/model/ascii/con_frac_train.asc", 
  agri_frac = 
    "~/model/ascii/agr_frac_train.asc", 
  water_frac = 
    "~/model/ascii/water_frac_train.asc",
  decid_frac = 
    "~/model/ascii/dec_frac_train.asc",
  maize_frac =
    "~/model/ascii/maize_frac_train.asc",
  lowcover_frac = 
    "~/model/ascii/low_frac_train.asc",
  urban_frac = 
    "~/model/ascii/urban_frac_train.asc", 
  shannon_wiener = 
    "~/model/ascii/shannon_train.asc", 
  forest_patch_density = 
    "~/model/ascii/forest_dens_train.asc", 
  mean_forest_patch_size = 
    "~/model/ascii/forest_size_train.asc", 
  road_density = 
    "~/model/ascii/road_dens_train.asc"
    
))

newvar <- raster :: stack (c(
  other_agri_dist =
    "~/model/ascii/agr_d_pred.asc",
  con_dist = 
    "~/model/ascii/con_d_pred.asc",
  decid_dist = 
    "~/model/ascii/dec_d_pred.asc",
  lowcover_dist = 
    "~/model/ascii/low_d_pred.asc", 
  mais_dist = 
    "~/model/ascii/maize_d_pred.asc",
  water_dist = 
    "~/model/ascii/water_d_pred.asc", 
  conif_frac = 
    "~/model/ascii/con_frac_pred.asc", 
  agri_frac = 
    "~/model/ascii/agr_frac_pred.asc", 
  water_frac = 
    "~/model/ascii/water_frac_pred.asc",
  decid_frac = 
    "~/model/ascii/dec_frac_pred.asc",
  maize_frac =
    "~/model/ascii/maize_frac_pred.asc",
  lowcover_frac = 
    "~/model/ascii/low_frac_pred.asc",
  urban_frac = 
    "~/model/ascii/urban_frac_pred.asc", 
  shannon_wiener = 
    "~/model/ascii/shannon_pred.asc", 
  forest_patch_density = 
    "~/model/ascii/forest_dens_pred.asc", 
  mean_forest_patch_size = 
    "~/model/ascii/forest_size_pred.asc", 
  road_density = 
    "~/model/ascii/road_dens_pred.asc"
    
))

```

# 3 Preparation data 
```{r}
data_wild_boar <- read.csv("~/model/data_wild_boar_final.csv",
                 sep = ",",
                 header = TRUE)

pseudo_absences_shp <- st_read("~/model/pseudo_1102.shp")
data_wild_boar$wild_boar <- 1

pseudo_absences <- as.data.frame(pseudo_absences_shp)
colnames(pseudo_absences)[colnames(pseudo_absences) == "x"] <- "NEAR_X"
colnames(pseudo_absences)[colnames(pseudo_absences) == "y"] <- "NEAR_Y"
pseudo_absences$repetition1 <- TRUE
pseudo_absences <- pseudo_absences %>%
                                   select(NEAR_X,NEAR_Y,repetition1)
combined_data <- merge(data_wild_boar, pseudo_absences, by = c("NEAR_X", "NEAR_Y"), all = TRUE)
myPAtable <- data.frame(PA1 = ifelse(combined_data$wild_boar == 1, TRUE, FALSE)) 
myPAtable$PA1 <- ifelse(is.na(myPAtable$PA1 ), TRUE, myPAtable$PA1  ) 

```

# 4 Formatting data
```{r}
wild_boar_dataset_2 <- BIOMOD_FormatingData(
                                           resp.var = combined_data$wild_boar,
                                           expl.var = variables, 
                                           resp.xy = combined_data[, c("NEAR_X", "NEAR_Y")], 
                                           resp.name = "wild_boar", 
                                           PA.strategy = "user.defined",
                                           PA.user.table = myPAtable, 
                                           filter = TRUE
)
plot(wild_boar_dataset_2)
```

# 5 Individual statistical models
```{r}
options <- BIOMOD_ModelingOptions(
GLM = NULL,
GAM = NULL,
RF = NULL,
ANN = NULL,
MAXENT = list(path_to_maxent.jar = "C:/Users/lucre/Documents/model/maxent/maxent.jar") 
)
wild_boar_models2 <- BIOMOD_Modeling(bm.format = wild_boar_dataset_2, 
                                    models = c("MAXENT", "GLM", "GAM", "RF", "ANN"), 
                                    models.pa = NULL,
                                    bm.options = options,
                                    CV.strategy = "random",
                                    CV.nb.rep = 5,
                                    CV.perc = 0.70,
                                    var.import = 5,
                                    metric.eval = "ROC",
                                    do.full.models = TRUE)

```

# 6 Model performance
```{r}
scores_ind <- get_evaluations(wild_boar_models2)
bm_PlotEvalBoxplot(bm.out = wild_boar_models2, group.by = c('algo', 'algo'))
bm_PlotEvalBoxplot(bm.out = wild_boar_models2, dataset = "validation", group.by = c('algo', 'algo'))
```

# 7 Response plots individual statistical models
```{r}
bm_PlotResponseCurves(
  bm.out =wild_boar_models2,
  models.chosen = "all",
  new.env = get_formal_data(wild_boar_models2, "expl.var"),
  show.variables = get_formal_data(wild_boar_models2, "expl.var.names"), 
  do.bivariate = FALSE,
  fixed.var = "median"
)
```

# 8 variable importance individual models
```{r}
var <- bm_PlotVarImpBoxplot(
  bm.out =  wild_boar_models2,
  group.by = c( "expl.var", "run","algo"),
  do.plot = FALSE
)
var2 <- as.data.frame(var$tab)
```

# 9 current and new area projection individual statistical models
```{r}
# for the current region 
wild_boar_models_current <- BIOMOD_Projection(bm.mod = wild_boar_models2, 
                                              new.env = variables,
                                              proj.name = "current2", 
                                              models.chosen = "all"
                                              
  
)
plot(wild_boar_models_current)


wild_boar_models_new <- BIOMOD_Projection(bm.mod = wild_boar_models2, 
                                              new.env = newvar,
                                              proj.name = "new", 
                                              models.chosen = "all"
                                            
)
plot(wild_boar_models_new)
```

# 10 BIOMOD-ensemble modelling 
```{r}
wild_boar_ensemble <- BIOMOD_EnsembleModeling(
  bm.mod = wild_boar_models2,
  em.by = "all",
  em.algo = "EMwmean",
  models.chosen = "all", 
  metric.select = "ROC", 
  var.import = 5, 
  metric.select.thresh = 0.6, 
  EMwmean.decay = 'proportional', 
  metric.select.dataset = "validation"
)

```
# 11 variable importance and evaluation scores
```{r}
var <- bm_PlotVarImpBoxplot(
  bm.out =  wild_boar_ensemble,
  group.by = c( "expl.var", "algo","full.name"),
  do.plot = FALSE
)
var3 <- as.data.frame(var$tab)
scores <- get_evaluations(wild_boar_ensemble)

```

# 12 Response plots ensemble model
```{r}
bm_PlotResponseCurves(
  bm.out =wild_boar_ensemble,
  models.chosen = "all",
  new.env = get_formal_data(wild_boar_models2, "expl.var"),
  show.variables = get_formal_data(wild_boar_models2, "expl.var.names"), 
  do.bivariate = FALSE,
  fixed.var = "median"
)

```

# 13 current and new area projection ensemble model
```{r}
projection_current <- BIOMOD_EnsembleForecasting(bm.em = wild_boar_ensemble,
                                             bm.proj = wild_boar_models_current,
                                             models.chosen = "all"
                                             )
plot(projection_current)
projection_newvar <- BIOMOD_EnsembleForecasting(bm.em = wild_boar_ensemble,
                                             bm.proj = wild_boar_models_new,
                                             models.chosen = "all"
                                             )
```




